Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-07-28 review-responses.tex

  \begin{point}{} 
    I think it would be helpful to say a bit more about the vaquita earlier in the manuscript. Although some of the authors have worked on the vaquita recently, it is still a fairly obscure organism, 
    so it would be worth describing its unfortunate situation and why it is an interesting comparison with humans throughout these simulations.  
\end{point}

  \reply{
      Good idea; we have added a bit more context, at \revref and \llname{vaquita_note}.
  }

  \begin{point}{} % 
    It seems as though everything on the backend is being done by SLiM in the demonstrations here 
    (unless I'm missing something about how the sweeps are done), so the runtimes are not really the product of the effort of the authors here. 
    Still, I think it would be good to provide a bit more information on the runtimes for prospective users to get a sense of what doing selection in stdpopsim entails. 
    (It's mentioned that the sweeps were time-consuming, but details aren't given as far as I can tell.)  
\end{point}

  \reply{
    We have now included two additional figures showing the runtimes for both the whole genome simulations (used in demographic and DFE inference) and the 5Mb segment simulations of human chromosome 1 (used in the analyses of sweep detection) at \revref and \llname{sim-sweeps-runtime}.
    With respect to the sweep simulations, we show that the simulation of sweeps in each of the segments takes approximately 3 minutes \llname{sim-sweeps-runtime}.
    Given we simulated 1000 replicates of each one of the 100 segments, this means that the total runtime for the sweep simulations is approximately 300,000 minutes, or 5,000 hours.
    However, whole chromosome 1 simulations with BGS and sweeps would have been much more time consuming, each taking approximately 200 minutes \revref (or in total, 20,000,000 minutes, or 333,333 hours).

### 2025-07-29 review-responses.tex

  %%%%%%%%%%%%%%
  \reviewersection{2}
  \begin{quote}
    \color{gray}
    In this manuscript, Gower et al. extend the functionality of stdpopsim to include selection. 
    They test the performance of a few different methods that infer historical population sizes, 
    the distribution of fitness effects of new deleterious mutations, and those that detect selective sweeps using the simulations in SLiM. 
    The added functionality will be useful to the users of stdpopsim.
\end{quote}

  Thanks for your constructive comments on our paper.

  \begin{point}{\revref} % \&75
    How do you deal with overlapping annotations in a gff?
  \end{point}

  \reply{
    We have added a note to the implementation section about how we handle overlapping annotations in a GFF file.
    In short, we do not allow for overlapping annotations in a GFF file, and we prioritize the annotation being modeled (e.g., CDS) 
    over potentially overlapping annotations of other types (e.g., exons). 
    Further, we do not currently support multiple overlapping annotations of different types.
    \revref.
  }

### 2025-07-29 analysis2_manuscript.tex
feat: added details about overlapping annotations and response

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    To extend \stdpopsim's functionality to include models of selection,
    we introduce two new classes
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects (DFE) of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where
    each mutation type is associated with a dominance coefficient,
    \revpoint{1}{2}
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    %IG: I moved this mention of h-s relationship up to here with a forward ref to newly added methods section
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep[][see \textbf{Methods}]{huber2018gene}.
    %by defining ranges of selection coefficients and associating each range with a different dominance coefficient. 
    %IG: I moved this label to the newly added methods section --> \revpoint{1}{5}
    Each \texttt{Annotation} object represents a collection of functional
    genomic elements (e.g., exons, introns, intergenic regions, etc.)
    whose coordinates are specified using a GFF3 annotation file. \revpoint{2}{1}
    More concretely, consider implementing a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
    with their proportions set to their expected prevalence in exons.
    The neutral mutation type is associated with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).

>>>>

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    To extend \stdpopsim's functionality to include models of selection,
    we introduce two new classes
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects (DFE) of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where
    each mutation type is associated with a dominance coefficient,
    \revpoint{1}{2}
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    %IG: I moved this mention of h-s relationship up to here with a forward ref to newly added methods section
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep[][see \textbf{Methods}]{huber2018gene}.
    %by defining ranges of selection coefficients and associating each range with a different dominance coefficient. 
    %IG: I moved this label to the newly added methods section --> \revpoint{1}{5}
    Each \texttt{Annotation} object represents a collection of functional
    genomic elements (e.g., exons, introns, intergenic regions, etc.)
    whose coordinates are specified using a GFF3 annotation file. 
    For the purposes of \stdpopsim, we do not allow for overlapping annotations -- 
    genomic elements of the given annotation being modeled (e.g., CDS) are prioritized over overlapping annotations of other types (e.g., exons)
    and we do not currently support multiple overlapping annotations of different types. \revpoint{2}{1}
    More concretely, consider implementing a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
    with their proportions set to their expected prevalence in exons.
    The neutral mutation type is associated with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
